RadLib: a radiative property model library for CFD

### 2021-10-06 manuscript/manuscript.tex
VBS clarity/style revisions---introduction

Modeling radiative heat transfer in participating media in high temperature systems is a challenging problem. Combustion is one of the key radiative transfer applications, and combustion CFD is uniquely challenging because it involves numerous tightly coupled physical phenomena---including multicomponent mass transfer, convective and radiative heat transfer, turbulent fluid dynamics, multi-phase and particle flows, and complex chemical reaction kinetics---that span many orders of magnitude in both length and time scales. Combustion simulations are often complicated by a lack of accurate radiation modeling, which can be difficult to implement and computationally expensive relative to overall simulation cost. Frequently, radiative heat transfer modeling in combustion systems is either neglected or oversimplified, resulting in significant errors flame temperatures~\citep{Modest_2016}. Additionally, some late-stage flame phenomena, particularly soot behavior and flame sheet breakthrough, increasingly depend on and influence the magnitude of radiative heat losses and are difficult to simulate accurately without robust radiation modeling \cite{Modest_2013}. These problems motivated development of the RadLib library, which implements three radiation property models with C++, Python, and Fortran interfaces and provides a framework for additional models and further development, giving researchers convenient access to advanced radiation property modeling tools.

The core mechanisms of radiative phenomena are physically and mathematically complex. Unlike other modes of heat transfer, radiation is governed by a set of integro-differential equations that depend on both direction and wavelength, known as the radiative transfer equations (RTE). Computational radiation solvers require two distinct but interconnected parts: (1) a solution approach for the radiative transfer equations (RTE) and (2) a radiative property model.
Complexity, implementation, and computational cost of these two parts can vary widely depending on the models chosen. The optically thin model for solution of the RTE is trivial, while the discrete ordinates model (DOM) and reverse Monte Carlo ray tracing (RMCRC) model, among others, are significantly more involved. Similarly, radiation property models range from simple correlations for computing mean absorption coefficients to gray gas models, including the weighted sum of gray gases (WSGG) model, to more advanced spectral line-weighted-sum-of-gray-gases (SLW) models to full line-by-line (LBL) treatment. In practical engineering simulations, the solution of the RTE is often the most complex and costly part of radiative simulation, but advanced property models such as the Rank Correlated SLW (RCSLW), discussed here, are nontrivial models to understand, implement, and validate.
RadLib addresses this second radiative property model part of radiation simulation by providing a set of implemented and validated radiation property models that can be applied to various solution approaches for the RTE.

For an absorbing, emitting, and non-scattering medium (considered here), the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
%
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$.  Solving the RTE requires information about the radiative properties of the materials and media in the system.  Radiative property data typically takes the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model.  Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules and depend on the local thermochemical state of the system (e.g., temperature, pressure, and composition).
Radiation property models are typically developed as correlations from high-resolution spectral databases, reducing the huge amount of spectral data contained in those databases and simplifying the evaluation of absorption coefficients \cite{Zhang_2002b}.

The direct solution of the spectral RTE using so-called line-by-line calculations, in which the RTE is solved for each individual spectral line, is computationally prohibitive for most practical configurations and is used primarily for model validation. Some band models rely on estimated or averaged "line-of-sight" property values that cannot be directly incorporated into the RTE.
%dol
Modern band models require high-resolution databases and reorder emission levels rather than averaging across sequential absorption bands, but these can lose accuracy when applied to nonhomogeneous media and may still require large numbers of RTE evaluations \citep{Modest_2016}.

Global models are an important class of radiation property models that use spectrally-integrated radiation properties \cite{Hottel_1967,Modest_2013,Modest_2016}. Assuming spectrally uniform (gray) properties over a wavenumber region $j$, we can integrate Eq.~\ref{e:RTE_general} over the fixed wavenumber ranges $\eta_j$ to obtain
%
\begin{linenomath}
\begin{equation} \label{e:RTE_wsgg}
    \frac{dI_j}{ds} = -\kappa_jI_j + a_j\kappa_jI_b, \hspace{20pt} j=0,1,...,n,
\end{equation}
\end{linenomath}
%
where $I_b$ is the blackbody intensity, $I_j$ is the radiative intensity in the region defined by gray gas $j$, $\kappa_j$ is the spectrally integrated absorption coefficient for gray gas $j$, and $a_j$ is a weighting factor defined by
%
\begin{linenomath}
\begin{align}
    &a_j= \frac{\int_{\eta_j}I_{b,\eta}d\eta}{I_b},\\
    &\sum_j a_j = 1.
\end{align}
\end{linenomath}
%
Given spatial profiles of $\kappa_j$ and $a_j$, Eq.~\ref{e:RTE_wsgg} can be evaluated for each of the $n$ gray gases. Normally, an additional ``clear gas'', $j=0$, is included that represents transparent windows in the spectrum for which $\kappa_0=0$. This gray gas approach allows calculation of the total radiative intensity $I$ by summing the individual gray gas:
%
\begin{linenomath}
\begin{equation}
I=\int_{\eta=0}^{\infty}I_{\eta}d\eta=\sum_{j=1}^n I_j.
\end{equation}
\end{linenomath}
%

The heat flux $\mathbf{q}$ (in units of kW/m$^2$) and volumetric radiative heat source $Q$ (kW/m$^3$) can be calculated by integrating over directional solid angles ($\mathbf{s}\Omega$) on the sphere and finding the divergence of the resulting vector field,
%
\begin{linenomath}
\begin{equation}
    \mathbf{q} = \int_{4\pi}I(\mathbf{s})\mathbf{s}d\Omega.
\end{equation}
\end{linenomath}
%
\begin{linenomath}
\begin{equation}
    Q = -\nabla\cdot\mathbf{q}.
\end{equation}
\end{linenomath}
%
Global models are especially attractive in computational heat transfer applications because they can provide good accuracy while limiting computational costs.

>>>>

Modeling radiative heat transfer in participating media in high temperature systems is a challenging problem. Combustion is one of the key radiative transfer applications, and combustion CFD is uniquely challenging because it involves numerous tightly coupled physical phenomena---including multicomponent mass transfer, convective and radiative heat transfer, turbulent fluid dynamics, multi-phase and particle flows, and complex chemical reaction kinetics---that span many orders of magnitude in both length and time scales.
Accurate radiation modeling often complicates combustion simulations because it can be difficult to implement and computationally expensive relative to the overall simulation cost.
%Combustion simulations are often complicated by a lack of accurate radiation modeling, which can be difficult to implement and computationally expensive relative to overall simulation cost.
Radiative heat transfer modeling in combustion systems is frequently neglected or oversimplified, resulting in significant errors flame temperatures~\citep{Modest_2016}.
%Frequently, radiative heat transfer modeling in combustion systems is either neglected or oversimplified, resulting in significant errors flame temperatures~\citep{Modest_2016}.
Additionally, some late-stage flame phenomena, particularly soot behavior and flame sheet breakthrough, increasingly depend on and influence the magnitude of radiative heat losses and are difficult to simulate accurately without robust radiation modeling \cite{Modest_2013}. These problems motivated development of the RadLib library, which implements three radiation property models with C++, Python, and Fortran interfaces and provides a framework for additional models and further development, giving researchers convenient access to advanced radiation property modeling tools.

The core mechanisms of radiative phenomena are physically and mathematically complex.
Unlike other modes of heat transfer, radiation is governed by the radiative transfer equations (RTE), a set of integro-differential equations that depend on both direction and wavelength.
%Unlike other modes of heat transfer, radiation is governed by a set of integro-differential equations that depend on both direction and wavelength, known as the radiative transfer equations (RTE).
Computational radiation solvers require two distinct but interconnected parts: (1) a solution approach for the radiative transfer equations (RTE) and (2) a radiative property model. The complexity, implementation, and computational cost of these two parts can vary widely depending on the models chosen.
For example, the optically thin approximation for solution of the RTE is trivial, while the discrete ordinates model (DOM) and reverse Monte Carlo ray tracing (RMCRC) model, among others, are significantly more involved.
%The optically thin model for solution of the RTE is trivial, while the discrete ordinates model (DOM) and reverse Monte Carlo ray tracing (RMCRC) model, among others, are significantly more involved.
Similarly, radiation property models range from simple correlations for computing mean absorption coefficients to gray gas models, including the weighted sum of gray gases (WSGG) model, to more advanced spectral line-weighted-sum-of-gray-gases (SLW) models to full line-by-line (LBL) treatment.
In practical engineering simulations, the solution of the RTE is often the most complex and costly part of radiative simulation, but advanced property models such as the Rank Correlated SLW (RCSLW) model, discussed in Section~\ref{s:RCSLW}, are nontrivial to understand, implement, and validate.
%In practical engineering simulations, the solution of the RTE is often the most complex and costly part of radiative simulation, but advanced property models such as the Rank Correlated SLW (RCSLW), discussed here, are nontrivial models to understand, implement, and validate.
RadLib addresses the second part of the radiation solver requirements by providing a set of implemented and validated radiation property models that can be applied alongside various solution approaches for the RTE.
%RadLib addresses this second radiative property model part of radiation simulation by providing a set of implemented and validated radiation property models that can be applied to various solution approaches for the RTE.

For an absorbing, emitting, and non-scattering medium (considered here), the RTE solved for the intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
%For an absorbing, emitting, and non-scattering medium (considered here), the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
%
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the spectral radiative intensity, $I_{b,\eta}$ is the blackbody spectral radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$.
%where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$.
Solving the RTE requires information about the radiative properties of the materials and media in the system. Radiative property data typically take the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model. Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules and depend on the local thermochemical state of the system (e.g., temperature, pressure, and composition).
Radiation property models are typically developed as correlations from high-resolution spectral databases, reducing the huge amount of spectral data contained in those databases and simplifying the evaluation of absorption coefficients \cite{Zhang_2002b}.

The direct solution of the spectral RTE using so-called line-by-line (LBL) calculations, in which the RTE is solved for each individual spectral line, is computationally prohibitive for most practical configurations and used primarily for model validation.
%The direct solution of the spectral RTE using so-called line-by-line calculations, in which the RTE is solved for each individual spectral line, is computationally prohibitive for most practical configurations and is used primarily for model validation.
Simple band models rely on estimated or averaged line-of-sight property values that cannot be directly incorporated into the RTE.
%Some band models rely on estimated or averaged "line-of-sight" property values that cannot be directly incorporated into the RTE.
%dol
Modern band models require high-resolution databases and reorder emission levels rather than averaging across sequential absorption bands, but these can lose accuracy when applied to nonhomogeneous media and may still require large numbers of RTE evaluations \citep{Modest_2016}.

Global models are an important class of radiation property models that use spectrally-integrated radiation properties \cite{Hottel_1967,Modest_2013,Modest_2016}.
Assuming spectrally uniform (gray) properties over a wavenumber region $j$, we can integrate Eq.~\ref{e:RTE_general} over a set of fixed wavenumber ranges $\eta_j$ to obtain
%Assuming spectrally uniform (gray) properties over a wavenumber region $j$, we can integrate Eq.~\ref{e:RTE_general} over the fixed wavenumber ranges $\eta_j$ to obtain
%
\begin{linenomath}
\begin{equation} \label{e:RTE_wsgg}
    \frac{dI_j}{ds} = -\kappa_jI_j + a_j\kappa_jI_b, \hspace{20pt} j=0,1,...,n,
\end{equation}
\end{linenomath}
%
where $I_b$ is the total blackbody intensity, $I_j$ is the total radiative intensity in the region defined by gray gas $j$, $\kappa_j$ is the spectrally integrated absorption coefficient for gray gas $j$, and $a_j$ is a weighting factor defined by
%where $I_b$ is the blackbody intensity, $I_j$ is the radiative intensity in the region defined by gray gas $j$, $\kappa_j$ is the spectrally integrated absorption coefficient for gray gas $j$, and $a_j$ is a weighting factor defined by
%
\begin{linenomath}
\begin{align}
    &a_j= \frac{\int_{\eta_j}I_{b,\eta}d\eta}{I_b},\\
    &\sum_j a_j = 1.
\end{align}
\end{linenomath}
%
Given spatial profiles of $\kappa_j$ and $a_j$, Eq.~\ref{e:RTE_wsgg} can be evaluated for each of the $n$ gray gases. Normally, an additional ``clear gas'', $j=0$, is included to represent transparent windows in the spectrum for which $\kappa_0=0$.
%Given spatial profiles of $\kappa_j$ and $a_j$, Eq.~\ref{e:RTE_wsgg} can be evaluated for each of the $n$ gray gases. Normally, an additional ``clear gas'', $j=0$, is included that represents transparent windows in the spectrum for which $\kappa_0=0$.
This gray gas approach allows calculation of the total radiative intensity $I$ by summing the individual gray gas intensities:
%This gray gas approach allows calculation of the total radiative intensity $I$ by summing the individual gray gas:
%
\begin{linenomath}
\begin{equation}
I=\int_{\eta=0}^{\infty}I_{\eta}d\eta=\sum_{j=1}^n I_j.
\end{equation}
\end{linenomath}
%

The heat flux $\mathbf{q}$ (in units of kW/m$^2$) and volumetric radiative heat source $Q$ (kW/m$^3$) can be calculated by integrating over directional solid angles ($\mathbf{s}\Omega$) on the sphere and finding the divergence of the resulting vector field, respectively:
%The heat flux $\mathbf{q}$ (in units of kW/m$^2$) and volumetric radiative heat source $Q$ (kW/m$^3$) can be calculated by integrating over directional solid angles ($\mathbf{s}\Omega$) on the sphere and finding the divergence of the resulting vector field,
%
\begin{linenomath}
\begin{equation}
    \mathbf{q} = \int_{4\pi}I(\mathbf{s})\mathbf{s}d\Omega.
\end{equation}
\end{linenomath}
%
\begin{linenomath}
\begin{equation}
    Q = -\nabla\cdot\mathbf{q}.
\end{equation}
\end{linenomath}
%
Global models are especially attractive in computational heat transfer applications because they can provide good accuracy while limiting computational costs.
